Sky Tomography for 3D Aerosol Distribution Recovery

### 2013-07-27 article.tex
bookmark

\noindent {\bf Extinction}: Sun rays (SR) irradiate a small volume
that includes particles of a certain type.  Each particle has an {\em
  extinction cross section} $\sigma$ for
interacting with the irradiance. The density of the particles is $n$. Per unit volume, the {\em extinction coefficient} is $\beta= \sigma n$.
The volume has infinitesimal length $dl$. Then, the relative portion
of extinct SR irradiance is the unitless differential optical
  depth, $d\tau= \beta dl=\sigma n dl$.
The {\em optical depth} aggregates in extended propagation:
\begin{align}
  \tau= \int d\tau=\int \beta dl=\int \sigma n dl \;.
  \label{eq:tau}
\end{align}
Through an attenuating volume, the {\em transmittance} exponentially
decays with the optical depth:
\begin{align}
  t=\exp(-\tau)
  \label{eq:beer-lambert}
\end{align}

\noindent {\bf Scattering}: Interaction of a single particle with the
irradiance is by absorption and scattering. The weight of scattering
(to all directions), relative to the total extinction is given by the
unitless {\em single scattering albedo} $\varpi$ of the particle. The
{\em scattering coefficient} $\alpha$ is
\begin{align}
  \alpha= \varpi\beta=\varpi \sigma n \;.
  \label{eq:alph}
\end{align}
The angular distribution of scattering is determined by a {\em phase
  function} $P$, which is normalized: its integral over all solid
angles is unit. Part of the light scatters towards a camera's LOS, as
illustrated in \cref{fig:groundgrid}. The angle between the SR and LOS
is the scattering angle $\Phi^{\rm scatter}$. The {\em angular scattering coefficient}
$\tilde\alpha$ is
\begin{align}
  \tilde\alpha(\Phi^{\rm scatter}) = \varpi\beta P(\Phi^{\rm scatter})
  = \varpi \sigma n P(\Phi^{\rm scatter}).
  \label{eq:alphabasic}
\end{align}

The phase function is often approximated by a parametric expression,
as the Henyey-Greenstein~\cite{Cornette1995} function,
\begin{align}
  P_g(\Phi^{\rm scatter})\approx \sfrac{3}{8\pi} (1 - g^2)(1+\mu^2)
  (2+g^2)^{-1}(1 + g^2 - 2g\mu)^{\sfrac{-3}{2}}
  \label{eq:aerosol_scatter}
\end{align}
where $g$ is an {\em anisotropy parameter} and $\mu\equiv \cos \Phi^{\rm scatter}$.
\\

\noindent {\bf Air molecules}: Scattering by air molecules follows the
{\em Rayleigh} model. The phase function is
\begin{align}
  P^{\rm air}\left[\mu\right] &\approx \sfrac{3}{16\pi}(1+\mu^2)
  \label{eq:rayleighP}
\end{align}
and the single scattering albedo is $\varpi^{\rm air}$=1. Air molecular density $n^{\rm air}$ falls
off approximately exponentially with altitude, with a
characteristic~\cite{Levi1980} falloff height $H^\mathrm{air}=8\
\si{\km}$. Thus~\cite{Levi1980}, the coefficients for
extinction and scattering by air molecules can be modeled by
\begin{align}
  \alpha^{\rm air}(h, \lambda)=\beta^{\rm air}(h, \lambda) \approxeq
  1.09 \times 10^{-3}\lambda^{-4}
  \exp(-h/H^\mathrm{air})
  \;,%\left[-\frac{h}{H^\mathrm{air}}\right].
  \label{eq:rayleighbeta}
\end{align}
where $\lambda$ is the wavelength, in microns.

@@

Consider a set of ground-based wide-angle cameras
% ~\cite{Cossairt2011, Peleg01omnistereo:panoramic,
% Koppal:2011:WMS:2191740.2192152}
looking upwards (\cref{fig:groundgrid}). The set performs {\em
  integral imaging}~\cite{integral} of the sky. Per viewpoint, the
view azimuth and elevation angles (relative to the zenith and north)
are encapsulated in the vector ${\bm{\Theta}}$. Any camera $c$ is at a
distinct fixed location ${\bm x}$, capturing raw image
$i_c({\bm{\Theta}})$. This paper takes the first step, of formulating
a principle of recovering the 3D atmospheric aerosols distribution,
using such data. To show feasibility, we perform radiative transfer
using two different forward models, enabling cross-validation.  A
forward model takes as input the 3D aerosol distribution, and outputs
images as if captured at various viewpoints. One model uses Monte
Carlo (MC) methods (\cref{sec:monte-carlo-simul}): it is stochastic,
slow but naturally expresses multiple-scattering. Hence, this approach
simulates realistic scenes of arbitrary complexity.  The second model
uses the single-Scattering approximation
(\cref{sec:single-scatt-model}). It is less accurate than MC, but
derives an analytic closed form solution to the radiative
transfer. This closed form enables simple inversion of the model, to
estimate the underlying atmosphere (\cref{sec:inverse-problem}).

@@

Any voxel generally contains both air molecules and aerosols.  We
assume that within any voxel, the molecular parameters $\{ \beta^{\rm
  air}(k),\alpha^{\rm air}(k)\}$ and the aerosol parameters $\{
\sigma^{\rm aerosol}(k),\varpi^{\rm aerosol}(k), n^{\rm
  aerosol}(k),g(k)\}$ are constants, e.g., corresponding to the value
at each voxel center.  Following \cref{eq:tau}, the optical depth
between the TOA and voxel $k$ is
\begin{equation}
  \tau_{\rm SR}(k)= \sum_{q \in[{\rm SR},k]} \beta(q) l_{\rm SR}(q) \;\;.
  \label{eq:lSRn}
\end{equation}
We can write \cref{eq:lSRn} using matrix notation, $\vect{\tau}_{\rm
  SR}=\OpDistance^{\mathrm{Sun \rightarrow voxel}} \vect{\beta}$,
where $\vect{\tau}_{\rm SR}$ and $\vect{\beta}$ are column stack
vector representations of $\tau_{\rm SR}(k)$ and $\beta(q)$,
respectively. Similarly, the optical depth along a LOS to voxel $k$ is
expressed by element $k$ of $\vect{\tau}_{{\rm LOS}_c} =
\OpDistance^{\mathrm{voxel \rightarrow cam}} \vect{\beta}$ Here
$\vect{\tau}_{{\rm LOS}_c}$ is a column stack vector representation of
$\tau_{{\rm LOS}_c}(k)$.  The SR and LOS paths joining at each voxel
can form a single path from the TOA to camera $c$, yielding
\begin{equation}
  {\bm D}_c=
  {\bm D}^{\mathrm{Sun \rightarrow voxel}}+
  {\bm D}^{\mathrm{voxel \rightarrow cam}}_c
  \quad\text{ and }\quad
  \vect{\tau}_c=
  \vect{\tau}_{\rm SR}+
  \vect{\tau}_{{\rm LOS}_c}
  \;\;.
  \label{eq:D_tau_total}
\end{equation}

@@

Combining \cref{eq:tautotal,eq:pc,eq:IcP}, the captured image is thus
\begin{align}
  {\bm i}_c= L^{\rm TOA} {\vect{\Pi}}_c \left\{
    (\tilde{\vect{\alpha}}^{\rm air}_c + \tilde{\vect{\alpha}}^{\rm
      aerosol}_c) \odot \exp\{-[\vect{\tau}_c^{\rm air} + {\bm D}_c
    (\vect{\sigma} \odot {\bm n})]\} \right\}
  \label{eq:bigIc}
\end{align}

To simplify the case in study, we focus on situations in which there is only a single
type of aerosol over a scene. Hence, the three-element vector
$[{\sigma}^{\rm aerosol},\varpi^{\rm aerosol},g]$ is uniform across
the scene, but the density distribution ${\bm n}$ is variable. Then,
\cref{eq:bigIc} degenerates to
\begin{align}
  {\bm i}_c= L^{\rm TOA} {\vect{\Pi}}_c \Big\{
  \exp\{-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm aerosol} {\bm D}_c
  {\bm n})\}\odot [\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
    aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
    scatter}_c) \odot{\bm n} ] \Big\} .
  \label{eq:bigIA}
\end{align}

@@

In general terms, the variables are the set \mbox{$\left\{{\bm
      n},\left\{\vect{\varpi}_{\lambda},\vect{\sigma}_{\lambda},{\bm
        g}_{\lambda}\right\}_{\lambda}\right\}$} where $\lambda$
denotes the spectral (color) channel.  There are $N_{\rm spectral}$
spectral channels.  Thus, the number of unknown variables is in
principle \mbox{$N_{\rm voxels}(1+3N_{\rm spectral})$}.

Prior knowledge reduces the unknowns. For simplicity we assume a
single aerosol type dominates a scene. Furthermore, in the following
section we assume that we know the type of the particle
\cite{Martonchik2009}. Hence, we focus here on retrieving ${\bm n}$.

The recovery task is phrased as optimization of an objective function,
that fits the image-formation model (\cref{eq:bigIc,eq:bigIA}) to
$N_{\rm views}$ measured photographs $\{ {\bm i}^{\rm
  measured}_c\}_{c=1}^{N_{\rm views}}$:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bm n} =
  \argmin_{{\bm n}} F({\bm n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bm n})
  = \sum_{c=1}^{N_{\rm views}}
  \| \Mask \roundy{{\bm i}^{\rm measured}_c - {\bm i}_c({\bm n})}\|^2_2 + \eta \cdot \Psi({\bm n})
\end{equation}
Here $\Psi({\bm n})$ is a regularization term that expresses some
prior knowledge on the solution and $\eta$ is the regularization
coefficient.  We use $\Mask$ as a weight map to mask the influence of
the sun and the horizon. These regions are prone to dominant multiple
scattering.  One common measurement of smoothness are the second order
derivatives.  These can be calculated using the \emph{Laplacian}
operator, $\nabla^2$.  We enforce smoothness of the solution by
minimizing the signal energy of the \emph{Laplacian} of the signal
\begin{equation}
  \label{eq:regularizer}
  \Psi({\bm n}) = \| \nabla^2{\bm n}\|^2_2 = \| \Laplacian{\bm n}\|^2_2
\end{equation}
where $\Laplacian$ is a matrix representation of the \emph{Laplacian}
operator.

We solve~\cref{eq:minobjectiveA} using standard optimization tools.
The gradient of $F$ with respect to ${\bm n}$ is
\begin{align}
  \Grad{{\bm n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bm J}_{{\bm i}_c}({\bm n})\right]} [{\bm i}^{\rm
    measured}_c - {\bm i}_c({\bm n})] + 2 \cdot \eta \cdot
  \Laplacian^T\Laplacian{\bm n} \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bm J}_{{\bm i}_c}({\bm n})$ is the Jacobian of the
vector ${\bm i}_c$ with respect to ${\bm n}$.  Element $(\Theta,k)$ of
this matrix differentiates the intensity in pixel ${\bm \Theta}$ (in
viewpoint $c$) with variation of the density at voxel $k$, i.e.,
$\partial i_c({\bm \Theta})/\partial{n^{\rm aerosol}(k)}$.  A
close-form expression of our Jacobian and its detailed derivation is
given in \cref{sec:gradient-derivation}.

>>>>

\noindent {\bf Extinction}: Sun rays (SR) irradiate a small volume
that includes particles of a certain type.
Each particle has an {\em extinction cross section} for
interacting with the irradiance. For an aerosol the cross section is denoted $\sigma^{\rm aerosol}$.
The aerosol density is $n$. Per unit volume, the {\em extinction coefficient} due to aerosols is
$\beta^{\rm aerosol}= \sigma^{\rm aerosol} n$. In addition, the atmosphere contains air molecules.
The extinction coefficient due to the molecules is $\beta^{\rm air}$.
The volume has infinitesimal length $dl$. Then, the relative portion
of extinct SR irradiance is the unitless differential optical
  depth, $d\tau= (\beta^{\rm aerosol}+\beta^{\rm air}) dl$.
The {\em optical depth} aggregates in extended propagation:
\begin{align}
  \tau= \int d\tau=\int (\beta^{\rm aerosol}+\beta^{\rm air}) dl
       =\int (\sigma^{\rm aerosol} n +\beta^{\rm air}) dl
       = \tau^{\rm air}+ \int \sigma^{\rm aerosol} n dl \;,
  \label{eq:tau}
\end{align}
where $\tau^{\rm air}=\int \beta^{\rm air} dl$.
Through an attenuating atmosphere, the {\em transmittance} exponentially
decays with the optical depth:
\begin{align}
  t=\exp(-\tau) \;.
  \label{eq:beer-lambert}
\end{align}

\noindent {\bf Scattering}: Interaction of a single particle with the
irradiance is by absorption and scattering. The weight of scattering
(to all directions), relative to the total extinction is given by the
unitless {\em single scattering albedo} of the particle. For an aerosol,
the single scattering albedo is denoted $\varpi^{\rm aerosol}$.
The {\em scattering coefficient} due to aerosols in the volume is
\begin{align}
  \alpha^{\rm aerosol}= \varpi^{\rm aerosol}\beta^{\rm aerosol}
                       =\varpi^{\rm aerosol} \sigma^{\rm aerosol} n \;.
  \label{eq:alph}
\end{align}
The angular distribution of scattering by an aerosol is determined by a {\em phase function} $P^{\rm aerosol}$, which is normalized: its integral over all solid angles is unit. Part of the light scatters towards a camera's LOS, as
illustrated in \cref{fig:groundgrid}. The angle between the SR and LOS
is the scattering angle $\Phi^{\rm scatter}$. The {\em angular scattering coefficient} due to aerosols is
\begin{align}
  \tilde\alpha^{\rm aerosol}(\Phi^{\rm scatter}) 
  =  \alpha^{\rm aerosol} P^{\rm aerosol}(\Phi^{\rm scatter})
  = \varpi^{\rm aerosol} \sigma^{\rm aerosol} n 
        P^{\rm aerosol}(\Phi^{\rm scatter}).
  \label{eq:alphabasic}
\end{align}
The phase function is often approximated by a parametric expression,
as the Henyey-Greenstein~\cite{Cornette1995} function,
\begin{align}
  P^{\rm aerosol}_g(\Phi^{\rm scatter})\approx \sfrac{3}{8\pi} (1 - g^2)(1+\mu^2)
  (2+g^2)^{-1}(1 + g^2 - 2g\mu)^{\sfrac{-3}{2}}
  \label{eq:aerosol_scatter}
\end{align}
where $g$ is an {\em anisotropy parameter} and $\mu\equiv \cos \Phi^{\rm scatter}$.

Scattering by air molecules follows the
{\em Rayleigh} model. The phase function is
\begin{align}
  P^{\rm air}\left[\mu\right] &\approx \sfrac{3}{16\pi}(1+\mu^2)
  \label{eq:rayleighP}
\end{align}
and the single scattering albedo in visible light is $\varpi^{\rm air}$=1. Air molecular density falls off approximately exponentially with altitude $h$, with a characteristic~\cite{Levi1980} falloff height $H^\mathrm{air}=8\
\si{\km}$. Thus~\cite{Levi1980}, the coefficients for
extinction and scattering by air molecules can be modeled by
\begin{align}
  \alpha^{\rm air}(h, \lambda)=\beta^{\rm air}(h, \lambda) \approxeq
  1.09 \times 10^{-3}\lambda^{-4}
  \exp(-h/H^\mathrm{air})
  \;,%\left[-\frac{h}{H^\mathrm{air}}\right].
  \label{eq:rayleighbeta}
\end{align}
where $\lambda$ is the wavelength, in microns.

@@

Consider a set of ground-based wide-angle cameras
% ~\cite{Cossairt2011, Peleg01omnistereo:panoramic,
% Koppal:2011:WMS:2191740.2192152}
looking upwards (\cref{fig:groundgrid}). The set performs {\em
  integral imaging}~\cite{integral} of the sky. Per viewpoint, the
view azimuth and elevation angles (relative to the zenith and north)
are encapsulated in vector ${\bm{\Theta}}$. Any camera $c$ is at a
distinct fixed location ${\bm x}$, capturing raw image
$i_c({\bm{\Theta}})$. This paper takes the first step, of formulating
a principle of estimating the 3D atmospheric aerosols distribution,
using such data. To show feasibility, we perform radiative transfer
using two different forward models, enabling cross-validation.  A
forward model takes as input the 3D aerosol distribution, and outputs
images as if captured at various viewpoints. One model uses Monte
Carlo (MC) methods (\cref{sec:monte-carlo-simul}): it is stochastic,
slow but naturally expresses multiple-scattering. Hence, this approach
simulates realistic scenes of arbitrary complexity.  The second model
uses the single-scattering approximation
(\cref{sec:single-scatt-model}). It is less accurate than MC, but
solves radiative transfer in an analytic, closed form. This form enables simple inversion of the model, to
estimate the underlying atmosphere (\cref{sec:inverse-problem}).

@@

Any voxel generally contains both air molecules and aerosols.  We
assume that within any voxel, the molecular parameters $\{ \beta^{\rm
  air}(k),\alpha^{\rm air}(k)\}$ and the aerosol parameters $\{
\sigma^{\rm aerosol}(k),\varpi^{\rm aerosol}(k), n(k),g(k)\}$ are constants, e.g., corresponding to the value
at each voxel center.  Following \cref{eq:tau}, the optical depth
between the TOA and voxel $k$ is
\begin{equation}
  \tau_{\rm SR}(k)= \sum_{q \in[{\rm SR},k]} \beta(q) l_{\rm SR}(q) \;\;.
  \label{eq:lSRn}
\end{equation}
We can write \cref{eq:lSRn} using matrix notation, $\vect{\tau}_{\rm
  SR}=\OpDistance^{\mathrm{Sun \rightarrow voxel}} \vect{\beta}$,
where $\vect{\tau}_{\rm SR}$ and $\vect{\beta}$ are column stack
vector representations of $\tau_{\rm SR}(k)$ and $\beta(q)$,
respectively. Similarly, the optical depth along a LOS to voxel $k$ is
expressed by element $k$ of $\vect{\tau}_{{\rm LOS}_c} =
\OpDistance^{\mathrm{voxel \rightarrow cam}} \vect{\beta}$ Here
$\vect{\tau}_{{\rm LOS}_c}$ is a column stack vector representation of
$\tau_{{\rm LOS}_c}(k)$.  The SR and LOS paths joining at each voxel
can form a single path from the TOA to camera $c$, yielding
\begin{equation}
  {\bm D}_c=
  {\bm D}^{\mathrm{Sun \rightarrow voxel}}+
  {\bm D}^{\mathrm{voxel \rightarrow cam}}_c
  \quad\text{ and }\quad
  \vect{\tau}_c=
  \vect{\tau}_{\rm SR}+
  \vect{\tau}_{{\rm LOS}_c}
  \;\;.
  \label{eq:D_tau_total}
\end{equation}

@@

Combining \cref{eq:tautotal,eq:pc,eq:IcP}, the captured image is thus
\begin{align}
  {\bm i}_c= L^{\rm TOA} {\vect{\Pi}}_c \left\{
    (\tilde{\vect{\alpha}}^{\rm air}_c + \tilde{\vect{\alpha}}^{\rm
      aerosol}_c) \odot \exp\{-[\vect{\tau}_c^{\rm air} + {\bm D}_c
    (\vect{\sigma} \odot {\bm n})]\} \right\}
  \label{eq:bigIc}
\end{align}
We focus on situations where there is only a single
type of aerosol over a scene. Hence, the three-element vector
$[{\sigma}^{\rm aerosol},\varpi^{\rm aerosol},g]$ is uniform across
the scene, but the density distribution ${\bm n}$ is variable. Then,
\cref{eq:bigIc} degenerates to
\begin{align}
  {\bm i}_c= L^{\rm TOA} {\vect{\Pi}}_c \Big\{
  [\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
    aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
    scatter}_c) \odot{\bm n} ] 
  \odot   
  \exp[-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm aerosol} {\bm D}_c{\bm n})]
    \Big\} .
  \label{eq:bigIA}
\end{align}

@@

We seek to estimate the aerosol distribution. Most generally, this is the set 
$\left\{{\bm n},\left\{\vect{\varpi}_{\lambda},\vect{\sigma}_{\lambda},{\bm
        g}_{\lambda}\right\}_{\lambda}\right\}$
where some of the parameters depend on the spectral channel $\lambda$. Usually, however, a single aerosol type dominates a scene. Furthermore, we assume that the type of the aerosol above an area is known~\cite{Martonchik2009}. Hence, we focus on estimating the density distribution ${\bm n}$.

The recovery task is phrased as optimization of an objective function,
that fits the image-formation model (\cref{eq:bigIc,eq:bigIA}) to
$N_{\rm views}$ measured photographs $\{ {\bm i}^{\rm
  measured}_c\}_{c=1}^{N_{\rm views}}$:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bm n} =
  \argmin_{{\bm n}} F({\bm n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bm n})
  = \sum_{c=1}^{N_{\rm views}}
  \| \Mask \roundy{{\bm i}^{\rm measured}_c - {\bm i}_c({\bm n})}\|^2_2 
  + \eta \Psi({\bm n})
\end{equation}
where $\Mask$ is a map that masks the areas around 
the sun, which are overwhelmingly bright. %and the horizon.
%These regions are prone to dominant multiple scattering. 
In Eq.~(\ref{eq:objectiveA}), $\Psi({\bm n})$ is a regularization term that expresses some prior knowledge on the distribution, while $\eta$ is a regularization weight.  Since aerosol distributions are usually fuzzy, 
it can be beneficial that regularization be a {\em smoothness term}, which penalized energy in second order spatial derivatives (3D Laplacian). Hende, we use
\begin{equation}
  \label{eq:regularizer}
  \Psi({\bm n}) = \| \nabla^2{\bm n}\|^2_2 = \| \Laplacian{\bm n}\|^2_2
\end{equation}
where $\Laplacian$ is a matrix representation of the 3D \emph{Laplacian}
operator.

We solve~\cref{eq:minobjectiveA} using standard optimization tools.
The gradient of $F$ with respect to ${\bm n}$ is
\begin{align}
  \Grad{{\bm n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bm J}_{{\bm i}_c}({\bm n})\right]} [{\bm i}^{\rm
    measured}_c - {\bm i}_c({\bm n})] + 2 \eta \Laplacian^T\Laplacian{\bm n} \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bm J}_{{\bm i}_c}({\bm n})$ is the Jacobian of the
vector ${\bm i}_c$ with respect to ${\bm n}$.  Element $(\Theta,k)$ of
this matrix differentiates the intensity in pixel ${\bm \Theta}$ (in
viewpoint $c$) with variation of the aerosol density at voxel $k$, i.e.,
$\partial i_c({\bm \Theta})/\partial{n(k)}$.  A
close-form expression of the Jacobian and its detailed derivation is
given in \cref{sec:gradient-derivation}.
